---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "50%",
  fig.align = "center"
)
```

Starting with **rasterpic** is very easy! You just need a image (`png`, 
`jpeg/jpg` or `tif/tiff`) and a spatial object (from the **sf** or the
**terra**) package to start using it

## Basic usage

We use here as an example the shape of Austria:

```{r setup}
library(sf)
library(terra)
library(rasterpic)

x <- st_read(system.file("gpkg/austria.gpkg",
  package = "rasterpic"
),
quiet = TRUE
)
img <- system.file("img/vertical.png",
  package = "rasterpic"
)

# Create!

default <- rasterpic_img(x, img)

plotRGB(default)
plot(st_geometry(x), add = TRUE, col = "grey90")
```


## Options

The function provides several options for expanding, alignment and cropping.

### Expand

With this option the image is zoomed out of the spatial object:

```{r}
expand <- rasterpic_img(x, img, expand = 1)

plotRGB(expand)
plot(st_geometry(x), add = TRUE, col = "grey90")
```


### Alignment

Decide where to align the image:

```{r}
bottom <- rasterpic_img(x, img, valign = 0)

plotRGB(bottom)
plot(st_geometry(x), add = TRUE, col = "grey90")
```

### Crop and mask

Create impressive maps!:

```{r}
mask <- rasterpic_img(x, img, crop = TRUE, mask = TRUE)

plotRGB(mask)

maskinverse <- rasterpic_img(x, img,
  crop = TRUE,
  mask = TRUE,
  inverse = TRUE
)

plotRGB(maskinverse)
```
