---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  out.width = "100%"
)
```

Starting with **rasterpic** is very easy! You just need a image (`png`,
`jpeg/jpg` or `tif/tiff`) and a spatial object from the **sf** or the **terra**
package to start using it.

## Basic usage

We use here as an example the shape of Austria:

```{r setup}
library(sf)
library(terra)
library(rasterpic)

# Plot
library(tidyterra)
library(ggplot2)

# Shape and image
x <- read_sf(system.file("gpkg/austria.gpkg", package = "rasterpic"))
img <- system.file("img/vertical.png", package = "rasterpic")

# Create!

default <- rasterpic_img(x, img)

autoplot(default) +
  geom_sf(data = x)
```

## Options

The function provides several options for expanding, alignment and cropping.

### Expand

With this option the image is zoomed out of the spatial object:

```{r}
expand <- rasterpic_img(x, img, expand = 1)

autoplot(expand) +
  geom_sf(data = x)
```

### Alignment

Decide where to align the image:

```{r}
bottom <- rasterpic_img(x, img, valign = 0)

autoplot(bottom) +
  geom_sf(data = x)
```

### Crop and mask

Create impressive maps!:

```{r}
mask <- rasterpic_img(x, img, crop = TRUE, mask = TRUE)

autoplot(mask)

maskinverse <- rasterpic_img(x, img, crop = TRUE, mask = TRUE, inverse = TRUE)

autoplot(maskinverse)
```

## Objects admitted for geotagging

-   Spatial object of the **sf** package: `sf`, `sfc`, `sfg` or `bbox`.
-   Spatial objects of the **terra** package: `SpatRaster`, `SpatVector`,
    `SpatExtent`.
-   A vector of coordinates with the form `c(xmin, ymin, xmax, yman)`.

## Image formats admitted

**rasterpic** can parse the following image formats:

-   `png` files.
-   `jpg/jpeg` files.
-   `tif/tiff` files.
