---
title: "Get started"
description: Introduction to the rasterpic package.
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    warning: false
    message: false
    out.width: "100%"
---

Getting started with **rasterpic** is easy: you need an image (`png`,
`jpeg/jpg` or `tif/tiff`) and a spatial object from the **sf**, **terra** or
**stars** packages to begin.

## Basic usage

Here we use the shape of Austria as an example:

```{r}
#| label: fig-setup
#| fig-cap: Raster map geotagged with the coordinates of Austria
library(sf)
library(terra)
library(rasterpic)

# Plot.
library(tidyterra)
library(ggplot2)

# Set the shape and image.
x <- read_sf(system.file("gpkg/austria.gpkg", package = "rasterpic"))
img <- system.file("img/vertical.png", package = "rasterpic")

# Create the raster.

default <- rasterpic_img(x, img)

autoplot(default) +
  geom_sf(data = x)
```

## Options

`rasterpic_img()` provides several options for expansion, alignment and
cropping.

### Expand

With this option, the raster extent is expanded beyond the spatial object:

```{r}
#| label: fig-expand
#| fig-cap: Example image expansion
expand <- rasterpic_img(x, img, expand = 1)

autoplot(expand) +
  geom_sf(data = x)
```

### Alignment

Choose the alignment of the image within the spatial extent:

```{r}
#| label: fig-bottom
#| fig-cap: Example image alignment
bottom <- rasterpic_img(x, img, valign = 0)

autoplot(bottom) +
  geom_sf(data = x)
```

### Crop and mask

Crop and mask the image:

```{r}
#| label: fig-mask
#| fig-cap:
#|   - Example of masked image
#|   - Example of inverse masked image
mask <- rasterpic_img(x, img, crop = TRUE, mask = TRUE)

autoplot(mask)

maskinverse <- rasterpic_img(x, img, crop = TRUE, mask = TRUE, inverse = TRUE)

autoplot(maskinverse)
```

## Supported spatial objects for geotagging

- Spatial objects of the **sf** package: `sf`, `sfc`, `sfg` or `bbox`.
- Spatial objects of the **terra** package: `SpatRaster`, `SpatVector` and
  `SpatExtent`.
- Spatial objects of the **stars** package: `stars`.
- A numeric coordinate vector of the form `c(xmin, ymin, xmax, ymax)`.

`rasterpic_img()` is an S3 generic. The methods for extent-like inputs use the
object extent, and vector methods can also mask the image to the object shape.

## Supported image formats

**rasterpic** can parse the following image formats:

- `png` files.
- `jpg/jpeg` files.
- `tif/tiff` files.
