---
title: "Get started"
description: Introduction to the **rasterpic** package.
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    warning: false
    message: false
    out.width: "100%"
---

Getting started with **rasterpic** is straightforward: you need an image file
(`png`, `jpeg`/`jpg` or `tiff`/`tif`) and an object of a supported spatial class,
such as those from the **sf**, **terra** or **stars** packages.

## Basic usage

This example geotags an image using the shape of Austria:

```{r}
#| label: fig-setup
#| fig-alt: >
#|   Map with longitude on the horizontal axis and latitude on the vertical
#|   axis. A portrait photograph of a one-way sign is centered behind the gray
#|   silhouette of Austria. The photograph extends above and below the country
#|   to preserve its proportions.
#| fig-cap: Image geotagged with the coordinates of Austria.
library(sf)
library(terra)
library(rasterpic)

# Load plotting packages.
library(tidyterra)
library(ggplot2)

# Set the spatial object and image.
x <- read_sf(system.file("gpkg/austria.gpkg", package = "rasterpic"))
img <- system.file("img/vertical.png", package = "rasterpic")

# Geotag the image.
default <- rasterpic_img(x, img)

autoplot(default) +
  geom_sf(data = x)
```

## Options

`rasterpic_img()` provides options for expansion, alignment, cropping and
masking.

### Expansion

The `expand` argument expands the raster extent beyond the spatial object:

```{r}
#| label: fig-expand
#| fig-alt: >
#|   Map with longitude on the horizontal axis and latitude on the vertical
#|   axis. A portrait photograph of a one-way sign is centered behind the gray
#|   silhouette of Austria. The expanded photograph extends beyond the country
#|   on all four sides.
#| fig-cap: Example image expansion.
expand <- rasterpic_img(x, img, expand = 1)

autoplot(expand) +
  geom_sf(data = x)
```

### Alignment

The `halign` and `valign` arguments control the alignment of the image relative
to the spatial extent:

```{r}
#| label: fig-bottom
#| fig-alt: >
#|   Map with longitude on the horizontal axis and latitude on the vertical
#|   axis. A portrait photograph of a one-way sign sits behind the gray
#|   silhouette of Austria. The bottom edge of the photograph aligns with the
#|   southern edge of the country, leaving the extra image height above it.
#| fig-cap: Example image alignment.
bottom <- rasterpic_img(x, img, valign = 0)

autoplot(bottom) +
  geom_sf(data = x)
```

### Crop and mask

The `crop`, `mask` and `inverse` arguments control whether the raster is cropped
to the object's extent, masked to its shape or inversely masked:

```{r}
#| label: fig-mask
#| fig-alt:
#|   - >
#|     Map with longitude and latitude axes. A photograph of a one-way sign is
#|     visible only inside the silhouette of Austria, with the area outside the
#|     country left blank.
#|   - >
#|     Map with longitude and latitude axes. A photograph of a one-way sign
#|     fills the rectangular extent around Austria, while the country
#|     silhouette is blank. The inverse mask retains the image outside the
#|     country.
#| fig-cap:
#|   - Example masked image.
#|   - Example inversely masked image.
mask <- rasterpic_img(x, img, crop = TRUE, mask = TRUE)

autoplot(mask)

maskinverse <- rasterpic_img(x, img, crop = TRUE, mask = TRUE, inverse = TRUE)

autoplot(maskinverse)
```

## Supported spatial input classes

`rasterpic_img()` supports the following spatial input classes:

- **sf** classes: `sf`, `sfc`, `sfg` and `bbox`.
- **terra** classes: `SpatRaster`, `SpatVector` and `SpatExtent`.
- **stars** class: `stars`.
- A numeric coordinate vector of the form `c(xmin, ymin, xmax, ymax)`.

`rasterpic_img()` is an S3 generic. Methods for extent-like inputs use the
object's extent, and vector methods can also mask the image to its shape.

## Supported image formats

**rasterpic** can read the following image formats:

- PNG files (`png`).
- JPEG files (`jpeg`/`jpg`).
- TIFF files (`tiff`/`tif`).
