---
title: "Plotting terra rasters"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  out.width = "100%"
)
```

This article shows how to plot the results of `rasterpic_img()` with several
packages.

## Base plots

The most straightforward option is to use the base `plot()` methods provided by
the package **terra** (`terra::plotRGB()`)

```{r setup}

library(rasterpic)
library(terra)

# The flag of the United Kingdom
img <- system.file("img/UK_flag.png",
  package = "rasterpic"
)

uk <- sf::st_read(system.file("gpkg/UK.gpkg",
  package = "rasterpic"
),
quiet = TRUE
)

uk_img <- rasterpic_img(uk, img, mask = TRUE, inverse = TRUE)
plotRGB(uk_img)
```

## With ggplot2 + tidyterra

**tidyterra** provides full support for **terra** rasters:

```{r}

library(ggplot2)
library(tidyterra)

ggplot() +
  geom_spatraster_rgb(data = uk_img)
```

## With tmap

**tmap** can be also used to create great maps:

```{r}

library(tmap)

tm_shape(uk_img) +
  tm_graticules() +
  tm_rgb()
```

## With mapsf

**mapsf**, the replacement of **cartography**, also provides this functionality:

```{r}
library(mapsf)

mf_raster(uk_img)
mf_scale()
mf_inset_on(x = "worldmap", pos = "topright")
mf_worldmap(uk)
mf_inset_off()
```

## With maptiles

**maptiles** is a interesting package that provides the ability of downloading
map tiles from different providers. It also has a specific function for plotting
**terra** rasters:

```{r}

library(maptiles)

other_tile <- get_tiles(uk, crop = TRUE, zoom = 6)

other_tile_crop <- terra::crop(other_tile, uk_img)

plot_tiles(other_tile_crop)
plot_tiles(uk_img, add = TRUE)
```
