---
title: "Plotting terra rasters"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png",
  dpi = 300,
  out.width = "100%"
)
```

This article shows how to plot the `SpatRaster` produced by `rasterpic_img()`
with several packages.

## Base plots

The most straightforward option is to use the base `plot()` methods provided by
the package **terra** (`terra::plotRGB()`):

```{r setup, fig.alt="Example: plot with terra package"}
library(rasterpic)
library(terra)

# The flag of the United Kingdom
img <- system.file("img/UK_flag.png",
  package = "rasterpic"
)

uk <- sf::st_read(
  system.file("gpkg/UK.gpkg",
    package = "rasterpic"
  ),
  quiet = TRUE
)

uk_img <- rasterpic_img(uk, img, mask = TRUE, inverse = TRUE)
plotRGB(uk_img)
```

## With ggplot2 + tidyterra

**tidyterra** provides full support for **terra** `SpatRaster` objects:

```{r tterra, fig.alt="Example: plot with tidyterra package"}
library(ggplot2)
library(tidyterra)

ggplot() +
  geom_spatraster_rgb(data = uk_img)
```

## With tmap

**tmap** can also be used to create great maps:

```{r tmap, fig.alt="Example: plot with tmap package"}
library(tmap)

tm_shape(uk_img) +
  tm_graticules() +
  tm_rgb()
```

## With mapsf

**mapsf** also provides this functionality:

```{r mapsf, fig.alt="Example: plot with mapsf package"}
library(mapsf)

mf_raster(uk_img)
mf_scale()
mf_inset_on(x = "worldmap", pos = "topright")
mf_worldmap(uk)
mf_inset_off()
```

## With maptiles

**maptiles** is an interesting package that provides the ability to download map
tiles from different providers. It also has a specific function for plotting
**terra** `SpatRaster` objects:

```{r maptiles, fig.alt="Example: plot with maptiles package"}
library(maptiles)

other_tile <- get_tiles(uk, crop = TRUE, zoom = 6)

other_tile_crop <- terra::crop(other_tile, uk_img)

plot_tiles(other_tile_crop)
plot_tiles(uk_img, add = TRUE)
```

## References

```{r refs, echo=FALSE, results='asis'}
cits <- c(
  citation("tmap")[1], citation("mapsf")[1], citation("tidyterra")[1],
  citation("terra")[1], citation("ggplot2")[1]
)
print(cits, style = "html")
```

## Session info

<details>

<summary>Details</summary>

```{r session, echo=FALSE}
if (!require("sessioninfo")) {
  install.packages("sessioninfo")
}
sessioninfo::session_info()
```

</details>
