---
title: "Plotting SpatRaster objects"
description: Several ways to plot geotagged SpatRaster objects in R.
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    warning: false
    message: false
    dpi: 300
    tidy: "styler"
    dev: "ragg_png"
    out.width: "100%"
---

This article shows several ways to plot the `SpatRaster` object returned by
`rasterpic_img()`.

## Base plots

The most direct option is to use plotting functions from the **terra** package,
such as `terra::plotRGB()`:

```{r}
#| label: fig-terra
#| fig-cap: "Plot with the **terra** package"
library(rasterpic)
library(terra)

# Use the flag of the United Kingdom.
img <- system.file("img/UK_flag.png",
  package = "rasterpic"
)

uk <- sf::st_read(
  system.file("gpkg/UK.gpkg",
    package = "rasterpic"
  ),
  quiet = TRUE
)

uk_img <- rasterpic_img(uk, img, mask = TRUE, inverse = TRUE)
plotRGB(uk_img)
```

## With ggplot2 and tidyterra

The **tidyterra** package provides **ggplot2** support for **terra**
`SpatRaster` objects:

```{r}
#| label: fig-tterra
#| fig-cap: "Plot with the **tidyterra** package"
library(ggplot2)
library(tidyterra)

ggplot() +
  geom_spatraster_rgb(data = uk_img)
```

## With tmap

The **tmap** package can also create maps from `SpatRaster` objects:

```{r}
#| label: fig-tmap
#| fig-cap: "Plot with the **tmap** package"
library(tmap)

tm_shape(uk_img) +
  tm_graticules() +
  tm_rgb()
```

## With mapsf

The **mapsf** package can also plot `SpatRaster` objects:

```{r}
#| label: fig-mapsf
#| fig-cap: "Plot with the **mapsf** package"
library(mapsf)

mf_raster(uk_img)
mf_scale()

mf_inset_on(x = "worldmap", pos = "topright")
mf_worldmap(uk)
mf_inset_off()
```

## With maptiles

The **maptiles** package can download map tiles from different providers. It
also provides functions for plotting **terra** `SpatRaster` objects:

```{r}
#| label: fig-maptiles
#| fig-cap: "Plot with the **maptiles** package"
library(maptiles)

other_tile <- get_tiles(uk, crop = TRUE, zoom = 6)

other_tile_crop <- terra::crop(other_tile, uk_img)

plot_tiles(other_tile_crop)
plot_tiles(uk_img, add = TRUE)
```

## References

```{r}
#| label: refs
#| echo: false
#| results: asis
cits <- c(
  citation("tmap")[1], citation("mapsf")[1], citation("tidyterra")[1],
  citation("terra")[1], citation("ggplot2")[1]
)
print(cits, style = "html")
```
